---
format: pdf
title: Abundância relativa de ocorrência de tartarugas marinhas a partir de avistagens em pontos fixos  
subtitle: Do conhecer ao preservar
author:
  - name: Letícia Verônica dos Santos
    email: leticia.veronica@outlook.com
    affiliations: 
        - id: some-tech
          name: Universidade Federal do Rio Grande - FURG
          department: Programa de Pós Graduação em Ambientometria
          address: Avenida Itália km8
          city: Rio Grande
          state: Rio Grande do Sul
          postal-code: 96203-535 
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
Atualmente existem sete espécies de tartarugas marinhas no mundo e apenas cinco com ocorrência no Brasil: a tartaruga-cabeçuda ("Caretta caretta"); a tartaruga-verde ("Chelonia mydas"); a tartaruga-de-pente italic("Eretmochelys imbricata"); a  tartaruga-de-couro ("Dermochelys coriacea") e a tartaruga-oliva ("Lepidochelys olivacea"). Tais espécies escolhem o litoral brasileiro para desova (ICMBIO), assim como para atividades de alimentação e ‘sociabilização’ (Caminho Marinho). A espécie "Chelonia mydas", é a que predomina nas amostragens em locais de alimentação no litoral sul do Brasil. O objetivo deste trabalho é "avaliar a variabilidade da abundância relativa de ocorrência de tartarugas marinhas na praia de Itapirubá-SC", buscando encontrar maiores ou menores probabilidades de ocorrência de acordo com parâmetros oceanográficos, ambientais, atmosféricos e sazonais. A análise utilizará o método de "Distribuição de Poisson", o qual irá avaliar a distribuição discreta de probabilidade aplicável a ocorrências de um número de eventos em um intervalo específico.

keywords: 
  - tartarugas marinhas1
  - "Chelonia mydas2"
  - litoral brasileiro3
  
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

<- model1a-num-names.bst
<- elsarticle-template-1a-num.tex

WOOD, Simon N. Generalized additive models: an introduction with R. 2006. 

ZUUR, A. F.; SAVELIEV, A. A.; IENO, E. N. A Begginner’s Guide to Generalised Additive Mixed Model with R. 2014

<-


Here are two sample references:  @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. 
You can sets extra options with `natbiboptions` variable in YAML header. Example 

```
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL 

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:

$$P(x=k)=e^{-u}*u^k  
/k!$$


$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {-}
